### 2024-07-29 Paper/paper2.tex

    Various characteristics of firm leadership such as gender, age, and background have been shown to be correlated with firm decisions and performance. I investigate this topic in the context of hospital executives with differing backgrounds: some with training as a medical doctor and some without. Hiring physician executives can be controversial. On one hand, physicians bring a unique combination of clinical and administration expertise, and they have insights that can potentially improve hospital outcomes (\cite{Stajduhar_2023}, \cite{Ahmed_2022}). Alternatively, physicians may not always be adequately trained to enter into high level management positions, and therefore worsen hospital outcomes (\cite{HarvardBusinessReview2018}). In this paper, I investigate how having clinically trained executives in leadership affects hospital response to pay-for-performance incentives.

    I scrape executive team names and titles from publicly available Tax Form 990s, filled out by nonprofit firms in the US. Combining this with publicly available hospital characteristics, I construct a novel data set of nonprofit hospitals from 2010-2014 containing information on both executive team characteristics and quality metrics. Nonprofit behavior in the hospital industry directly affects health care consumers, as private nonprofit hospitals make up 50\% of all hospitals in the US, and staff, on average, 207 beds. For comparison, for-profits make up 36\% of hospitals and staff 107 beds (\cite{ASPE_2023}).

    The selection of leaders makes it difficult to interpret any direct comparison of differing executive teams as causal. Therefore, I leverage an exogenous shock to hospital incentives, pay-for-performance policies enacted in the US in 2012, and compare the change in readmission and mortality rates of executive teams with and without clinical executives. Under the assumption that the two types of teams would have had similar trends in readmission and mortality rates absent the incentive change, I find that a lack of clinically trained executives leads a hospital to respond more drastically to pay-for-performance incentives.

    Having found that hospital leadership does, in fact, matter to hospital behavior, I investigate several mechanisms that might contribute to this difference. First, \textcolor{red}{the more drastic decrease in readmissions by nonclinical teams is not driven by decreasing average patient complexity. There is no difference in patient case mix between clinical and nonclinical executive teams}. Further, I comparing each type of leadership team with for-profit hospitals reveals that the absence of clinical executives leads a nonprofit hospital to respond more similarly to a for-profit hospital. That is, having clinical training on a leadership team leads a hospital to act less profit-driven than their general training counterparts. Finally, I show that changes in a hospital's propensity to hire a clinical executive are not correlated with the pay-for-performance policies. Thus, I exploit such changes in leadership teams to decompose the effect of clinical executives into signaling (clinical executives signal underlying hospital preferences), or managing (clinical executives manage the hospital differently). I find that the entire difference in response is driven by a managing effect.

@@

    Due to data limitations, many of these studies focus on large, publicly traded firms in the US. \citeauthor{brickley2010board} (\citeyear{brickley2010board}) has the only study to my knowledge answering this question in the US nonprofit context. Using exogenous variation in expected Medicare profits, the authors find that having an internal board of directors increases CEO compensation, and having physician board members decreases public donations (\cite{brickley2010board}). Additionally, two studies focus on hospital performance in other countries. \citeauthor{janke2019impact} (\citeyear{janke2019impact}) uses data from England to study whether CEOs affect hospital production, and find no association (\cite{janke2019impact}). However, \citeauthor{otero2022managers} (\citeyear{otero2022managers}) investigates the role of CEOs in public hospitals in Chile, and finds an 8\% decrease in mortality rates for hospitals with top managers. Further, a major contributor of this quality improvement is the exodus of older physicians as CEOs (\cite{otero2022managers}). I contribute to our knowledge of how executive team characteristics affect firm behaviors by (1) considering the important setting of nonprofit hospitals, which serve the majority of patients in the US, (2) considering a unique executive characteristic which is plausibly relevant, and (3) observing firm behavior in the presence of an exogenous incentive change.

@@

    Finally, I match each name with the National Plan and Provider Enumeration System (NPPES) database of all registered physicians. The goal is to match each physician with an NPI number in order to merge publicly available data on physician specialty, gender, and tenure. For any person who has a doctor title and uniquely matches the NPPES data, I record their NPI number. For anyone who has a doctor title and no matches, no doctor title and matches, or a doctor title and multiple matches, I use information on location and hospital name to manually match the NPI.

    For the purposes of this paper, I focus on executive teams as a whole instead of specific positions. The main reason for this is because the organization of teams varies so drastically across hospitals that a specific position, even if it exists in two different hospitals, can be meaningfully different. I drop anyone who is solely a board member, and focus on those who have executive in their title, or are labeled as a president or vice president of a department. I create various hospital-level characterizations based on their executive team, such as the existence of a clinically trained executive, the number of clinically trained executives, the number of total executives, and whether the hospital has a Chief Medical Officer. For my main sample, I limit to hospitals who do not change their propensity to hire a clinical executive for the duration of the sample. Summary statistics for these variables are shown in Table \ref{leader_sumstats}. On average, NFP hospitals in my sample have 4 executives and .4 clinically trained executives. Sixty-one percent of the sample never changes propensity to hire a clinically trained executive, and 33\% of NFPs in the sample have a Chief Medical Officer at some point.

     \import{Tables}{leader_sumstats.tex}

@@

    In Table \ref{tab:sumstats_samples}, I present a table of means of relevant variables for several hospital samples used throughout the analysis. First, there are 708 for-profit hospitals in the sample, shown in column (1). For-profit hospitals in the sample have 155 beds on average, 112-276 patients depending on the condition, and are highly likely to be affiliated with a system. Thirty six percent of for-profits were penalized for all three conditions at some point.

    \import{Tables}{sample_sumstats.tex}

    Columns 2-4 show variable means for not-for-profits in the sample, with all NFPs in column 2, NFPs that ever have a clinically trained executive in column 3, and NFPs that never have a clinically trained executive in column 4. NFPs are much less likely to be physician owned or system affiliated compared to FPs. They do, however, have similar bed size and case mix index to FPs. Twenty-two percent of all NFPs are penalized for all conditions. There are several differences among NFPs with and without clinically trained executives. Those with a clinically trained executive are larger in terms of beds and patients seen, are more likely to be an academic medical center, and see slightly more complex patients. Further, they are more likely to be penalized for every condition.

    Finally, columns 5-6 limit the sample of not-for-profits further using changes in executive teams. The NFPs included in this sample do not change their propensity to hire a clinically trained executive, they either always have the same number of clinically trained executives or always have no clinically trained executives. The differences in NFPs with and without clinically trained executives remains similar with this sample limitation. I discuss the implications of these differences for identification in Section \ref{sec:model}.

    I plot weighted average readmission and mortality rates across time for different hospital types in Figure \ref{fig:weighted_read_mort_graph}, where I adjust both rates by baseline patient complexity. Pay-for-performance incentive changes are shown as a dotted line between 2011 and 2012. For-profit hospitals have the highest readmission rates prior to 2012. While the readmission rate trend behaves similarly across hospital types prior to the programs, there are notable trend differences after the programs. For-profits decrease readmissions drastically. NFP trends depend on the presence of an executive MD, where those without a clinically trained executive respond drastically, more similarly to for-profits, than those with a clinically trained executive. In mortality rates, NFPs with a clinically trained executive have the lowest rates before the programs, but differentially continue with an increasing trend in mortality after the programs. While this is purely descriptive evidence, the trends in outcomes point to more similar behavior between for-profits and NFPs without a clinically trained executive in comparison to NFPs with a clinically trained executive.

    \begin{figure}[ht!]
    \centering
        \caption{Outcomes Over Time}
        \includegraphics[width=\textwidth]{Objects/weighted_read_mort_adjusted_graph.pdf}
        \label{fig:weighted_read_mort_graph}
    \end{figure}

@@

    My first goal is to estimate the effect of having clinical executive training on how hospitals change readmission and mortality rates in response to pay-for-performance incentives. Ultimately, I employ synthetic difference-in-differences, weighting more heavily non-clinical hospitals with readmission, mortality, and baseline case mix index that better match the clinical team trends. However, for clarity, I present the unweighted specification:

@@

     I rely on several assumptions to identify this effect. First, I assume that, absent the program enactments, hospitals with different leadership team characterization types would have behaved similarly in outcomes. While I see no evidence of pre-trends, Table \ref{tab:sumstats_samples} suggests that hospitals with a clinically trained executive are different than other hospitals along multiple characteristics. There may be concern that these differences between hospital types could be driving the results. Therefore, my main specification uses the synthetic difference-in-differences method established in \citeauthor{arkhangelsky2021synthetic} (\citeyear{arkhangelsky2021synthetic}). This method creates unit weights that more heavily factor comparison hospitals that are similar to the treat group hospitals prior to the programs. It also weighs time periods that balance pre-programs and post-programs for the comparison hospitals. Then, a simple two way fixed effects estimator is used with these weights, yielding a more robust average treatment effect that relies less on a strong parallel trends assumption (\cite{arkhangelsky2021synthetic}). I present the synthetic diff-in-diff estimates as the main results, and they are largely similar to the standard two way fixed effects results with no weights, shown in Appendix \ref{app:fullsample}. Further, the conceptual identification is the same as a standard difference-in-differences.

    \subsection{Results}

    The estimated difference in readmission rate trends for all hospital types in response to the programs are shown in Figure \ref{fig:clinicalsynthdid}. In panel \ref{fig:read_synth_clinical}, I compare readmission rate responses among nonprofit hospitals with and without clinical training on their executive team. Estimate $\beta$ is represented in the difference in slope after 2012 when pay-for-performance incentives began. Nonprofits without a clinically trained executive lowered readmission rates by .24 ppts more than those with a clinically trained executive.

    I also directly estimate the difference in mortality rates among NFPs with and without clinical experience in panel \ref{fig:mort_synth_clinical}. Again, the difference is not statistically significant, but the coefficient compared to the true nulls in panels \ref{fig:mort_synth_clinical} and \ref{fig:mort_synth_plotc} indicates NFPs without a clinical executive potentially decrease mortality more aggressively than NFPs with a clinical executive. In Appendix \ref{app:condition}, I show that these results are mainly driven by pneumonia patients.

@@

    \section{Are Clinical Executives less Profit Driven?}\label{sec:forprofit}

    A reason for the difference in hospital behavior with the presence of clinically trained executives could be how profit-driven a physician vs. a general leader is. To demonstrate this, consider hospital behavior under the enactment of pay-for-performance policies, where hospitals face a trade-off between profit and societal benefit. Pay-for-performance incentives essentially incorporate quality directly into a hospital's profit function (\cite{dranove2011health}), whereas a fee-for-service system only incentivizes quantity. A hospital that places all value on profit will only care about quality when it enters into the profit function, and therefore will respond more drastically to pay-for-performance than a hospital who already cared about quality through societal benefit before the incentives.\footnote{I present the theoretical model and mathematical results in Appendix \ref{sec:model}.} Therefore, observing hospital responses to pay-for-performance incentives reveals information about their underlying preferences.

    I compare readmission and mortality rates of each nonprofit leadership team type to for-profit hospitals. Since for-profit hospitals have revealed through their ownership status that they are largely profit-driven, comparing different executive team behavior to for-profit behavior helps reveal how profit driven certain leadership teams are. Thus, I estimate

@@

    \noindent for two subsets of data. First, I limit only to for-profits and nonprofits with clinical leadership. Second, I limit only to for-profits and nonprofits without clinical leadership. Thus, $\beta$ represents the effect of being for-profit compared to being nonprofit with a certain leadership team type.

    The identification assumptions I make here are largely similar to those in the main analysis, only now incorporate for-profit hospitals. First, I assume that, absent the change in incentives, for-profit and the relevant comparison group would have had parallel trends in outcomes. While this is not directly testable, both the synthetic difference in difference and event studies indicate parallel trends prior to the incentive changes. Second, I assume that the stability of nonprofit leadership teams is not correlated with the 2012 policy changes. I explore this assumption in Section \ref{sec:identification}. I also assume that there is no anticipation of the policy changes prior to 2012, which is reasonable as the rules of the policies were announced in October of 2011. Finally, I assume no other changes correlated with both hospital type (for-profit, nonprofit with and without clinically trained executives) and quality occurred in conjunction with the 2012 pay-for-performance policies.

@@

 \begin{figure}[ht!]
     \caption{Readmission Rate Synthetic Difference in Differences Results}
     \centering
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{For-Profit and NFP w/ MD Exec}
         \includegraphics[width=\textwidth]{Objects/read_fp_md_synth_graph.pdf}
         \label{fig:read_synth_plotb}
     \end{subfigure}%
     \vspace{5mm}
     \hfill
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{For-Profit and NFP w/out MD Exec}
         \includegraphics[width=\textwidth]{Objects/read_fp_nomd_synth_graph.pdf}
         \label{fig:read_synth_plotc}
     \end{subfigure}
        \label{fig:read_synth_plot}
    \end{figure}

    Next, I present estimated differences in mortality rates between different hospital types in Figure \ref{fig:mort_synth_plot}. In panel \ref{fig:mort_synth_plotb}, the estimated difference is between for-profits and nonprofits with clinically trained executives. The magnitude of the difference in mortality rates is .11, where only for-profits seem to be reacting to the incentive change. While still statistically insignificant, I argue that this still indicates a difference in behavior between the hospital types, particularly due to the documented noisiness of mortality measures (\cite{mackenzie2016measuring}). However, the estimated difference in mortality between for-profits and nonprofits without clinically trained executives is certainly zero.

    \begin{figure}[ht!]
     \caption{Mortality Rate Synthetic Difference in Differences Results}
     \centering
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{For-Profit and NFP w/ MD Exec}
         \includegraphics[width=\textwidth]{Objects/mort_fp_md_synth_graph.pdf}
         \label{fig:mort_synth_plotb}
     \end{subfigure}%
     \hfill
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{For-Profit and NFP w/out MD Exec}
         \includegraphics[width=\textwidth]{Objects/mort_fp_nomd_synth_graph.pdf}
         \label{fig:mort_synth_plotc}
     \end{subfigure}
        \label{fig:mort_synth_plot}
    \end{figure}

\section{Signaling vs. Managing Decomposition}

    To this point, I have limited the sample of nonprofit hospitals to those that don't change their propensity to hire a clinically trained executive over the sample period. However, under the assumption that hiring or firing these types of executives is not correlated with the pay-for-performance policy changes, these changes can be leveraged to understand further how executive characteristics affect hospital behavior. Particularly, a clinically trained executive could be a signal of the underlying objectives of the hospital, or hiring a clinically trained executive who carries out day-to-day operations in a distinct way effectively changes the objectives of the hospital.

@@

    I have discussed the assumptions of parallel trends, no anticipation, and no other confounding events in both Sections \ref{sec:clinical} and \ref{sec:forprofit}. These assumptions are also necessary here. In those Sections, I have discussed that stability of leadership team must not be correlated with the 2012 policy changes. Up to this point, that assumption was important because of the way I selected the sample. However, this decomposition depends directly on leveraging such leadership changes. Thus, I now explore this assumption in detail.

     I assess the validity of the exogenous changes assumption in two ways. First, I analyze whether changes in clinically trained executives occur more often in time periods of program enactment. Second, I analyze whether hospitals who end up getting penalized or receiving payments are more likely to make changes in clinical executive members than those who are not. That is, are hospitals that expect to be affected by the programs more likely to hire or fire clinically trained executives? I consider changes in having any clinically trained executive, as well as changes in the number of clinically trained executives. I estimate the following linear equations:

    \begin{equation}\label{eq:change1}
    \text{change}_{ht} = \sum_{j=2011}^{2014}\beta_j\mathbf{1}\{t=j\} + \alpha_h + \epsilon_{ht},
    \end{equation}

    \begin{equation}\label{eq:change2}
    \text{change}_{ht} = \beta(\text{Program Exposed}_{h} \times \text{Post Programs}_t)+ \alpha_h + \delta_t + \epsilon_{ht}.
    \end{equation}

    The variable $\text{Program Exposed}_{h}$ is an indicator for whether the hospital eventually was penalized under HRRP or received payments under HVBP. The estimates from this analysis are presented in Table \ref{tab:change_analysis}. Changes in any clinically trained executive are less likely to occur in 2012-2014 relative to 2010, indicating that executive teams are actually becoming more stable over time along the dimension of hiring clinically trained executives. Further, hospitals who end up being penalized through HRRP or receiving payments through HVBP are not more likely to change their propensity to hire any clinically trained executive. Results are similar when considering changes in the number of clinically trained executives. Thus, it seems unlikely that endogenous team formation is biasing the estimates. In Appendix \ref{app:changes}, I present various descriptive statistics of executive team changes over time, showing that hiring and firing of MD executives do not change much over time, the rate of clinical executives to total executives remains stable over time, and the distribution of timing of clinical executive changes.

     \import{Tables}{change_analysis.tex}

    The estimates of each decomposition specification are shown in Table \ref{tab:MD_noMD_readmort_decomp_synth}. Columns (1) and (2) show that the difference in readmission response is driven by having a clinically trained executive in 2012, compared to having a clinically trained executive at another point in time. Columns (3) and (4) support the same theory given the magnitudes of the coefficients, but are still statistically insignificant due to the noisiness of mortality. The results suggest that managers make a difference in behavior and objectives, rather than being a signal of already existing objectives.

>>>>

    Across firm types and industries, firm leader characteristics such as gender, age, and background have been shown to be correlated with behavior and performance. I investigate this topic in the context of hospital executives with differing backgrounds: some with training as a medical doctor and some without. Physicians bring a unique combination of clinical and administration expertise, and they have insights that can potentially improve hospital outcomes (\cite{Stajduhar_2023}, \cite{Ahmed_2022}). However, hiring physician executives can also be controversial, as physicians may not always be adequately trained to enter into high level management positions (\cite{HarvardBusinessReview2018}).

    In this paper, I investigate how having clinically trained executives in leadership affects quality of care in hospitals. I scrape executive team names and titles from publicly available Tax Form 990s, filled out by nonprofit firms in the US. Combining this with publicly available hospital characteristics, I construct a novel data set of nonprofit hospitals from 2010-2014 containing information on both executive team characteristics and quality metrics. Nonprofit behavior in the hospital industry directly affects health care consumers, as private nonprofit hospitals make up 50\% of all hospitals in the US, and staff, on average, 207 beds. For comparison, for-profits make up 36\% of hospitals and staff 107 beds (\cite{ASPE_2023}).

    The non-random selection of leaders makes it difficult to interpret any direct comparison of differing executive teams as causal. Therefore, I leverage an exogenous shock to hospital incentives, pay-for-performance policies enacted in the US in 2012, and compare the change in readmission and mortality rates of executive teams with and without clinical executives. Under the assumption that the two types of teams would have had similar trends in readmission and mortality rates absent the incentive change, I find that a lack of clinically trained executives leads a hospital to respond more drastically to pay-for-performance incentives. That is, while both types of hospital leadership teams decrease readmission rates after the policy change, those without clinical executives decrease readmissions at a faster rate than those with clinical executives.

    Having found that hospital leadership does, in fact, matter to hospital behavior, I investigate several mechanisms that might contribute to this difference. First, the more drastic decrease in readmissions by nonclinical teams is not driven by decreasing average patient complexity. Further, comparing each type of leadership team with for-profit hospitals reveals that the absence of clinical executives leads a nonprofit hospital to respond more similarly to a for-profit hospital. That is, having clinical training on a leadership team leads a hospital to act less profit-driven than their general training counterparts. Finally, I show that changes in a hospital's propensity to hire a clinical executive are not correlated with the pay-for-performance policies. Thus, I exploit such changes in leadership teams to decompose the effect of clinical executives into signaling (clinical executives signal underlying hospital preferences), or managing (clinical executives manage the hospital differently). I find that the entire difference in response is driven by a managing effect.

@@

    Due to data limitations, many of these studies focus on large, publicly traded firms in the US. \citeauthor{brickley2010board} (\citeyear{brickley2010board}) have the only study to my knowledge answering this question in the US nonprofit context. Using exogenous variation in expected Medicare profits, the authors find that having an internal board of directors increases CEO compensation, and having physician board members decreases public donations (\cite{brickley2010board}). Additionally, two studies focus on hospital performance in other countries. \citeauthor{janke2019impact} (\citeyear{janke2019impact}) use data from England to study whether CEOs affect hospital production, and find no association (\cite{janke2019impact}). However, \citeauthor{otero2022managers} (\citeyear{otero2022managers}) investigate the role of CEOs in public hospitals in Chile, and finds an 8\% decrease in mortality rates for hospitals with top managers. Further, a major contributor of this quality improvement is the exodus of older physicians as CEOs (\cite{otero2022managers}). I contribute to our knowledge of how executive team characteristics affect firm behaviors by (1) considering the important setting of nonprofit hospitals, which serve the majority of patients in the US, (2) considering a unique executive characteristic which is plausibly relevant, and (3) observing firm behavior in the presence of an exogenous incentive change.

@@

    Finally, I match each name with the National Plan and Provider Enumeration System (NPPES) database of all registered physicians. The goal is to match each physician with an NPI number in order to merge publicly available data on physician specialty, gender, and tenure. For any person who has a doctor title and uniquely matches the NPPES data, I record their NPI number. For anyone who has a doctor title and no matches, no doctor title and matches, or a doctor title and multiple matches, I use information on location and hospital name to manually match the NPI. I present a table of means for various characteristics of doctor executives in the data in Table \ref{doc_sumstats}. In total, there are 813 doctor executives. On average, they are 52 years old, and 11\% of them are female. While it is more likely they are a Chief Medical Officer than a Chief Executive Officer, still less than a third are CMOs. The most common specialty is internal medicine.

    \import{Tables}{doc_exec_stats_tab.tex}

    For the purposes of this paper, I focus on executive teams as a whole instead of specific positions. The main reason for this is because the organization of teams varies so drastically across hospitals that a specific position, even if it exists in two different hospitals, can be meaningfully different. I drop anyone who is solely a board member, and focus on those who have executive in their title, or are labeled as a president or vice president of a department. I create various hospital-level characterizations based on their executive team, such as the existence of a clinically trained executive, the number of clinically trained executives, the number of total executives, and whether the hospital has a Chief Medical Officer. For my main sample, I limit to hospitals who do not change their propensity to hire a clinical executive for the duration of the sample.

@@

    In Table \ref{tab:sumstats_samples_stable}, I present a table of means of relevant variables for several hospital samples used throughout the analysis. First, in column (1), are means for hospitals who ever have a clinical executive at any point in the sample. In column (2) are means for hospitals who have a clinical executive for the entire sample period, and column (3) shows means for hospitals who never have a clinical executive in the sample period.

    \import{Tables}{stable_sample_sumstats.tex}

    Hospitals that never have a clinical executive are smaller, less likely to be an academic medical center, and less likely to be affiliated with a system. Further, they have smaller executive teams on average by 1-2 people, and are less likely to have anyone acting as a CMO.\footnote{Though, they may title the position differently.} Finally, in regards to exposure to pay-for-performance programs, hospitals who never have clinical executives are less likely to receive benefits from the HVBP Program, but seem more similar to hospitals with clinical leadership in terms of penalties from HRRP.

    Next, I plot weighted average readmission, mortality rates, and patient case mix index across time for different hospital types in Figure \ref{fig:outcomes_graph}. Pay-for-performance incentive changes are shown as a dotted line in 2012. Average readmission rates trend similarly prior to 2012, and all hospital type exhibit a decrease in readmission after 2012. However, hospitals who consistently have a clinical executive show a slower rate of decrease in raw readmissions. When considering mortality, there is not a drastic change in 2012 for any hospital type, but hospitals who always have a clinical executive continue on an increasing trend in mortality where the other hospitals types remain more flat. Finally, there do not seem to be any differential trends in case mix index prior to or after 2012.

    \begin{figure}[ht!]
    \centering
        \caption{Outcomes Over Time}
        \includegraphics[width=.8\textwidth]{Objects/outcomes_graph.pdf}
        \label{fig:outcomes_graph}
    \end{figure}

@@

    First, I test empirically whether having clinical experience matters in how hospitals respond to the pay-for-performance policies effective in 2012. For a clear comparison, I limit the sample to hospitals who always have a clinical executive or never have a clinical executive. That is, these hospitals do not change their propensity to hire a clinical executive during the sample period. Therefore, an important assumption that I make is that this sample selection is not correlated with the program enactment. I investigate this assumption in Section \ref{sec:identification}, and find no evidence that hospitals are more or less likely to change their propensity to hire a clinical executive because of the pay-for-performance programs.

    I estimate a two way fixed effects specification, where the key variable is the interaction between having clinical executives and post 2012, when the programs began. Importantly, I am not estimating the effect of being penalized. All hospitals had incentive to improve quality as they are being compared to other hospitals who are also decreasing quality. Thus, I simply estimate the effect of being in a time period after the incentive changed for all hospitals. Ultimately, because there are differences in hospitals with different leadership types, I employ a synthetic difference-in-differences estimation strategy (\cite{arkhangelsky2021synthetic}). That is, hospitals with similar trends in readmission, mortality, and baseline case mix before the programs are more heavily weighted, and time periods that balance pre and post-program outcomes for hospitals without clinical executives. Below is the unweighted specification for clarity, where the identification assumptions are conceptually the same for the two specifications:

@@

    In the unweighted two way fixed effects specification, we have reason to believe that the parallel trends assumption may not hold due to observed differences in characteristics other than the leadership team, as discussed in Section \ref{sec:data}. Once incorporating the weights used in the synthetic differences in differences method, we ensure parallel trends prior to 2012. Thus, I assume that under the given clinical and nonclinical hospital composition, trends would have continued to be parallel absent the incentive change in 2012. A component of this assumption is the implicit assumption that no other event occurred in 2012 correlated with both the outcomes and composition of leadership team. \textcolor{red}{say something about ACA here? and need to add anticipation assumption}

    The estimated difference in readmission and mortality rate trends between hospitals with and without clinical executives are shown in Figure \ref{fig:clinicalsynthdid}. Estimate $\beta$ is represented in the difference in slope after 2012 when pay-for-performance incentives began. Panel \ref{read_synth_clinical} shows the estimated difference in readmission rates, and reveals that nonclinical teams decrease readmissions by .3 more than clinical teams. This is not a huge magnitude given an average readmission rate of 21, but reveals a meaningful difference in behavior given that both hospital types significantly decrease readmissions. In panel \ref{fig:mort_synth_clinical}, I show the estimated difference in mortality. While the magnitude of the difference is similar to that for readmission rates, mortality is a noisy measure, and I cannot rule out no difference in mortality response. In Appendix \ref{app:condition}, I show that these results are primarily driven by pneumonia patients.

@@

    \section{Selective Patient Practices}

    A method hospitals have been shown to employ in order to decrease readmission rates is being more selective with which patients to admit to the hospital. One hypothesis is that executive teams without any clinicians may be more willing to participate in such practices, whereas doctors prioritize care for all patients. This would explain why non-clinical hospitals are able to decrease readmissions more drastically.

    To test this, I employ case mix index as the outcome in specification \ref{eq:clinical}, to determine whether one type of hospital differentially changes patient composition after pay-for-performance incentives, where the same identification assumptions hold. I present the estimated difference in Figure \ref{fig:main_cmi_clinical}, which shows no differential response to the programs in terms of the complexity of patients admitted to the hospital. \textcolor{red}{sentence about the implications of this?}

    \begin{figure}[ht!]
        \centering
        \caption{Effect of Clinical Executive on Cases Mix Index}
        \includegraphics[width=0.5\textwidth]{Objects/cmi_md_nomd_synth_graph.pdf}
        \label{fig:main_cmi_clinical}
    \end{figure}

    \section{Does Specialty Matter?}

    \section{Are Clinical Executives Less Profit Driven?}\label{sec:forprofit}

    Another hypothesis that could explain the differential response due to clinical leadership is how profit-driven a clinician is compared to other training backgrounds such as an MBA. To demonstrate this, consider hospital behavior under the enactment of pay-for-performance policies, where hospitals face a trade-off between profit and societal benefit. Pay-for-performance incentives essentially incorporate quality directly into a hospital's profit function (\cite{dranove2011health}), whereas a fee-for-service system only incentivizes quantity. A hospital that places all value on profit will only care about quality when it enters into the profit function, and therefore will respond more drastically to pay-for-performance than a hospital who already cared about quality through societal benefit before the incentives.\footnote{I present the theoretical model and mathematical results in Appendix \ref{sec:model}.} Therefore, observing hospital responses to pay-for-performance incentives reveals information about their underlying preferences.

    I compare readmission and mortality rates of always clinical executives and never clinical executives to for-profit hospitals. While ideally I would also have leadership information for for-profit hospitals, the tax forms only have nonprofit hospitals. However, on average, for-profit hospitals have revealed through their ownership status that they heavily weigh profit. Thus, comparing different executive team behavior to for-profit behavior helps reveal how profit driven certain leadership teams are. Thus, I estimate

@@

    \noindent for two subsets of data. First, I limit only to for-profits and nonprofits with clinical leadership. Second, I limit only to for-profits and nonprofits without clinical leadership. Thus, $\beta$ represents the effect of being for-profit compared to being nonprofit with a certain leadership team type. I show summary statistics of relevant variables for for-profit hospitals in Appendix \textcolor{red}{add table}. Again, I employ a synthetic difference in differences strategy to make the parallel trends assumption more plausible.

    The identification assumptions I make here are largely similar to those in the main analysis, only now incorporate for-profit hospitals. First, I assume that, absent the change in incentives, for-profits and the relevant comparison group would have had parallel trends in outcomes. Second, I assume that the stability of nonprofit leadership teams is not correlated with the 2012 policy changes. I explore this assumption in Section \ref{sec:sig_man}. I also assume that there is no anticipation of the policy changes prior to 2012, which is reasonable as the rules of the policies were announced in October of 2011. Finally, I assume no other changes correlated with both hospital type (for-profit, nonprofit with and without clinically trained executives) and quality occurred in conjunction with the 2012 pay-for-performance policies.

@@

 \begin{figure}[ht!]
     \caption{Comparison to For-Profit: Readmission}
     \centering
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{For-Profit and Always Clinical Exec}
         \includegraphics[width=\textwidth]{Objects/fp_read_md_synth_graph.pdf}
         \label{fig:read_synth_plotb}
     \end{subfigure}%
     \vspace{5mm}
     \hfill
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{For-Profit and Never Clinical Exec}
         \includegraphics[width=\textwidth]{Objects/fp_read_nomd_synth_graph.pdf}
         \label{fig:read_synth_plotc}
     \end{subfigure}
        \label{fig:read_synth_plot}
    \end{figure}

    Next, I present estimated differences in mortality rates between different hospital types in Figure \ref{fig:mort_synth_plot}. In panel \ref{fig:mort_synth_plotb}, the estimated difference is between for-profits and nonprofits with clinically trained executives. The magnitude of the difference in mortality rates is .08, where clinical team hospitals continue on an increasing trend and for-profits remain stable in mortality rate. While I cannot rule out a lack of differential response, the documented noisiness of mortality measures (\cite{mackenzie2016measuring}) may mean that the true effect is different from zero. However, the estimated difference in mortality between for-profits and nonprofits without clinically trained executives is certainly zero.

    \begin{figure}[ht!]
     \caption{Mortality Rate Synthetic Difference in Differences Results}
     \centering
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{For-Profit and NFP w/ MD Exec}
         \includegraphics[width=\textwidth]{Objects/fp_mort_md_synth_graph.pdf}
         \label{fig:mort_synth_plotb}
     \end{subfigure}%
     \hfill
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{For-Profit and NFP w/out MD Exec}
         \includegraphics[width=\textwidth]{Objects/fp_mort_nomd_synth_graph.pdf}
         \label{fig:mort_synth_plotc}
     \end{subfigure}
        \label{fig:mort_synth_plot}
    \end{figure}

\section{Signaling vs. Managing Decomposition}\label{sec:sig_man}

    Given that having clinical leaders affects behavior, an important distinction is whether the existing clinical leaders \textit{reveal} the underlying objectives of the hospital, or \textit{manage} the hospital differently. Up to this point, I have limited the sample of nonprofit hospitals to those that don't change their propensity to hire a clinically trained executive over the sample period, which combines signalling and managing. However, under the assumption that hiring or firing these types of executives is not correlated with the pay-for-performance policy changes, such changes can be leveraged to decompose the effect into signaling vs. managing.

@@

    \subsection{Are Leadership Changes Correlated with Pay-for-Performance Policies?} \label{sec:changes}

    I have discussed the assumptions of parallel trends, no anticipation, and no other confounding events in both Sections \ref{sec:clinical} and \ref{sec:forprofit}. These assumptions are also necessary here. Further, this decomposition analysis depends heavily on leadership team changes not endogenously occurring due to pay-for-performance policies. Up to this point, that assumption was important because of the way I selected the sample. However, this decomposition depends directly on leveraging such leadership changes. Thus, I now explore this assumption in detail. First, I present raw means over time that represent the fraction of hospitals who hire a clinical executive, fire a clinical executive, or have any change in their propensity to have a clinical executive in Figure \ref{fig:change_means}. There do not seem to be any drastic changes in 2011 or 2012 when the policies were becoming relevant to hospitals.

    \begin{figure}[ht!]
    \centering
    \caption{Leadership Change Means Over Time}
    \includegraphics[width=0.8\textwidth]{Objects/change_means.pdf}
    \label{fig:change_means}
     \end{figure}

    Further, I assess whether this assumption is reasonable by analyzing how executive team changes are occurring over time for certain groups of hospitals. I estimate:

    \begin{equation}\label{eq:change2}
    \text{change}_{ht} = \sum_{j=2011}^{2014}\beta_j(\mathbf{1}\{t=j\}\times \text{Program Exposed})_{ht} + \alpha_h + \epsilon_{ht},
    \end{equation}

    where the variable $\text{Program Exposed}_{h}$ has three definitions. First, an indicator for whether the hospital was ever penalized under HRRP. Second, an indicator for whether the hospital ever received payments under HVBP. And finally, I leave out the Program Exposed variable to simply see how changes are correlated with given years for all hospitals. The outcome change$_{ht}$ is an indicator for whether the hospital changes the number of clinical executives in a given year. The estimates from this analysis are presented in Figure \ref{fig:change_analysis}. All estimates are not statistically different from zero, indicating that expectations of program exposure do not lead to changes in propensity to hire a clinical executive. Thus, it seems unlikely that endogenous team formation is biasing the estimates. In Appendix \ref{app:changes}, I present various descriptive statistics of executive team changes over time, showing that hiring and firing of MD executives do not change much over time, the rate of clinical executives to total executives remains stable over time, and the distribution of timing of clinical executive changes.

     \begin{figure}[ht!]
         \centering
         \caption{Leadership Analysis Results}
         \includegraphics[width=0.8\textwidth]{Objects/change_analysis_plot.pdf}
         \label{fig:change_analysis}
     \end{figure}

    \subsection{Decomposition Results}

    The estimates of each decomposition specification are shown in Table \ref{tab:MD_noMD_readmort_decomp_synth}. The decomposition for readmission rates is in Columns (1) and (2), and for mortality is in Columns (3) and (4). The first row compares ever clinical executive hospitals to never clinical executive hospitals, and the second row compares hospitals with a clinical executive in 2012 to hospitals who have a clinical executive in a year other than 2012. While there is no statistically significant estimate, the coefficients of the managing effect are very similar to the full effect found in Section \ref{sec:clinical}. The coefficients for the signaling effect are both statistically insignificant and close to zero in magnitude. Thus, I can rule out that clinical executives are simply a signal of underlying hospital objectives.
